A Formalization of Typed Lua

### 2015-08-18 doc/papers/dls15/paper.tex
fix a few typos in rules

\begin{figure*}[t]
{\footnotesize
\[
\begin{array}{c}
\begin{array}{c}
\mylabel{T-Constructor}\\
\dfrac{\begin{array}{c}
	\overline{\env_v, \penv \vdash e_1: F_1, \env_f}
	\;\;\;
\overline{\env_1, \penv \vdash e_2: F_2, \env_v}	\\
	T = \{\overline{F_1{:}vt(F_1,F_2)}\}_{unique} \;\;\;
	\env_{2} = merge(\overline{\env_{f}}) \;\;\; wf(T)
	\end{array}}
{\env_{1}, \penv \vdash \{\;\overline{[e_{1}] = e_{2}}\;\}:T, \env_{2}}
\end{array}
\begin{array}{c}
\mylabel{T-NewFieldUnique}\\
\dfrac{\begin{array}{c}
	\env_{1}(id) = \{ \overline{F:V} \}_{unique}\\
	\env_{1}, \penv \vdash e:F^{new}, \env_{2} \;\;\;
	\overline{F^{new} \not\lesssim F} \;\;\;
	V^{new} = vt(F^{new}, V)
	\end{array}}
{\begin{array}{l}
	\env_{1}, \penv \vdash id[e] {<}V{>}:V^{new},
	\env_{2}[id \mapsto \{ \overline{F{:}V}, F^{new}{:}V^{new}\}_{unique}]
	\end{array}}
\end{array}\\\\
\begin{array}{c}
\mylabel{T-IdRead}\\
\dfrac{\env(id) = F}
{\env, \penv \vdash id:close(F), \env[id \mapsto open(F)]}
\end{array}
\begin{array}{c}
\mylabel{T-IdWrite}\\
\dfrac{\env(id) = F}
{\env, \penv \vdash id_{l}:F, \env}
\end{array}
\begin{array}{c}
\mylabel{T-IndexIdRead}\\
\dfrac{\begin{array}{c}
	\env_{1}(id) = \{\overline{F:V}\} \;\;\;
	\env_{1}, \penv \vdash e_{2}:F^e, \env_{2} \;\;\;
	 F^e \lesssim F_{i}
	\end{array}}
{\env_{1}, \penv \vdash id[e_{2}]:rconst(V_{i}), \env_{2}}
\end{array}
\\ \\
\begin{array}{c}
\mylabel{T-IndexExpRead}\\
\dfrac{\begin{array}{c}
	\env_{1}, \penv \vdash e_{1}:\{\overline{F:V}\}, \env_{2}\\
	\env_{2}, \penv \vdash e_{2}:F^e, \env_{3} \;\;\;
	 F^e \lesssim F_{i}
	\end{array}}
{\env_{1}, \penv \vdash e_{1}[e_{2}]:rconst(V_{i}), \env_{3}}
\end{array}
\begin{array}{c}
\mylabel{T-CoerceClosed}\\
\dfrac{\env_{1}(id) \subtype F \;\;\; tag(F,closed)}
{\env, \penv \vdash {<}F{>} \; id:F, \env[id \mapsto reopen(F)]}
\end{array}
\begin{array}{c}
\mylabel{T-CoerceFixed}\\
\dfrac{\env_{1}(id) \subtype F \;\;\; tag(F,fixed)}
{\env, \penv \vdash {<}F{>} \; id:F, \env[id \mapsto F]}
\end{array}\\\\
\begin{array}{c}
\mylabel{T-Function}\\
\dfrac{\begin{array}{c}
	closeall(\env_{1})[\overline{id \mapsto F}], \penv[\ret \mapsto S] \vdash s, \env_{2}\\
	\env_{3} = openset(\env_{1}, frv(\mathbf{fun} \; (\overline{id{:}T}){:}S \; s)) \\
	\env_{4} = closeset(\env_{3}, fav(\mathbf{fun} \; (\overline{id{:}T}){:}S \; s))
	\end{array}}
{\env_{1}, \penv \vdash \mathbf{fun} \; (\overline{id{:}F}){:}S \; s:F_1 \times \ldots F_n \times \Void \rightarrow S, \env_{3}}
\end{array}
\begin{array}{c}
\mylabel{T-Local}\\
\dfrac{\begin{array}{c}
	\env_{1}, \penv \vdash el:E, \env_{2} \\
	E \lesssim F_1 \times \ldots \times F_n \times \Value{*} \\
	\env_{2}[\overline{id \mapsto F}], \penv \vdash s, \env_{3}
	\end{array}}
{\env_{1}, \penv \vdash \mathbf{local} \; \overline{id{:}F} = el \; \mathbf{in} \; s, (\env_{3} - \{\overline{id}\})[\overline{id \mapsto \env_2(id)}]}
\end{array}
\end{array}
\]
}
\caption{Typing Rules}
\label{fig:tablerules}
\end{figure*}

@@

Rule {\sc T-Constructor} also checks if the resulting type is {\em well-formed} with the predicate $wf$. Checking for well-formedness avoids ambiguous table types. An example of ambiguous table type is
$\{1:\Number, \Integer:\String, \Any:\Boolean\}$,
where indexing it with a key of type $\Integer$ could
yield either a number, a string or a boolean at runtime.

@@

\begin{figure*}[t]
{\footnotesize
\label{fig:projtypes}
\[
\begin{array}{c}
\begin{array}{c}
\mylabel{T-LocalProj}\\
\dfrac{\begin{array}{c}
	\env_{1}, \penv \vdash el:E, \env_{2}, (x,S)\\
	\env_{3} = \env_{2}[id_{1} \mapsto proj(E, 1), \ldots, id_{n} \mapsto proj(E, n)]\\
	\env_{3}, \penv[x \mapsto S] \vdash s, \env_{4} \;\;\;
	n = |\;\overline{id}\;|  
	\end{array}}
{\env_{1}, \penv \vdash \mathbf{local} \; \overline{id} = el \; \mathbf{in} \; s, (\env_{4}-\{\overline{id}\})[\overline{id \mapsto \env_2(id)}]}
\end{array}
\begin{array}{c}
\mylabel{T-ExpListProj}\\
\dfrac{\begin{array}{c}
	\env_{1}, \penv \vdash e_{i}:T_{i}, \env_{i+1} \;\;\;
	\env_{1}, \penv \vdash me:S, \env_{n+2} \;\;\; n=|\overline{e}|\\
	S = S_1 \sqcup S_2 \;\;\;m = max(|S_1|, |S_2|) \;\;\; x \notin \Pi\\
	\env_{f} = merge(\env_{1}, ..., \env_{n+2})
	\end{array}}
{\env_{1}, \penv \vdash \overline{e},me:}\\ {T_{1} \times ... \times T_{n} \times \pi_{1}^{x} \times ... \times \pi_{m}^{x}{*}, \env_{f}, (x,S)}
\end{array}\\\\
\begin{array}{c}
\mylabel{T-IfFilterProjThen}\\
\dfrac{\begin{array}{c}
	\env_{1}(id) = \pi_{i}^{x} \\ S_t = fonilpt(\Pi(x), i)\\ 
	finil(proj(\Pi(x), i)) = {\bf void} \\
	\env_{1}, \penv[x \mapsto S_t] \vdash s_{1}, \env_{2} 
	\end{array}}
{\env_{1}, \penv \vdash \mathbf{if} \; id \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}, \env_{2}}
\end{array}
\begin{array}{c}
\mylabel{T-IfFilterProjElse}\\
\dfrac{\begin{array}{c}
	\env_{1}(id) = \pi_{i}^{x} \\
	 S_e = finilpt(\Pi(x), i)\\
	fonil(proj(\Pi(x), i)) = {\bf void} \\
	\env_{1}, \penv[x \mapsto S_e] \vdash s_{2}, \env_{2} 
	\end{array}}
{\env_{1}, \penv \vdash \mathbf{if} \; id \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}, \env_{2}}
\end{array}
\begin{array}{c}
\mylabel{T-IfFilterProjBoth}\\
\dfrac{\begin{array}{c}
	\env_{1}(id) = \pi_{i}^{x} \\ S_t = fonilpt(\Pi(x), i) \; S_e = finilpt(\Pi(x), i)\\
	\env_{1}, \penv[x \mapsto S_t] \vdash s_{1}, \env_{2} \\
	\env_{1}, \penv[x \mapsto S_e] \vdash s_{2}, \env_{3} \\
	\env_{4} = join(\env_{2}, \env_{3}) 
	\end{array}}
{\env_{1}, \penv \vdash \mathbf{if} \; id \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}, \env_{4}}
\end{array}
\end{array}
\]}
\caption{Rules for Projection Types}
\end{figure*}

>>>>

\begin{figure*}[t]
{\footnotesize
\[
\begin{array}{c}
\begin{array}{c}
\mylabel{T-Constructor}\\
\dfrac{\begin{array}{c}
	\overline{\env_v, \penv \vdash e_1: F_1, \env_f}
	\;\;\;
\overline{\env_1, \penv \vdash e_2: F_2, \env_v}	\\
	T = \{\overline{F_1{:}vt(F_1,F_2)}\}_{unique} \;\;\;
	\env_{2} = merge(\overline{\env_{f}}) \;\;\; wf(T)
	\end{array}}
{\env_{1}, \penv \vdash \{\;\overline{[e_{1}] = e_{2}}\;\}:T, \env_{2}}
\end{array}
\begin{array}{c}
\mylabel{T-NewFieldUnique}\\
\dfrac{\begin{array}{c}
	\env_{1}(id) = \{ \overline{F:V} \}_{unique}\\
	\env_{1}, \penv \vdash e:F^{new}, \env_{2} \;\;\;
	\overline{F^{new} \not\lesssim F} \;\;\;
	V^{new} = vt(F^{new}, V)
	\end{array}}
{\begin{array}{l}
	\env_{1}, \penv \vdash id[e] {<}V{>}:V^{new},
	\env_{2}[id \mapsto \{ \overline{F{:}V}, F^{new}{:}V^{new}\}_{unique}]
	\end{array}}
\end{array}\\\\
\begin{array}{c}
\mylabel{T-IdRead}\\
\dfrac{\env(id) = F}
{\env, \penv \vdash id:close(F), \env[id \mapsto open(F)]}
\end{array}
\begin{array}{c}
\mylabel{T-IdWrite}\\
\dfrac{\env(id) = F}
{\env, \penv \vdash id_{l}:F, \env}
\end{array}
\begin{array}{c}
\mylabel{T-IndexIdRead}\\
\dfrac{\begin{array}{c}
	\env_{1}(id) = \{\overline{F:V}\} \;\;\;
	\env_{1}, \penv \vdash e_{2}:F^e, \env_{2} \;\;\;
	 F^e \lesssim F_{i}
	\end{array}}
{\env_{1}, \penv \vdash id[e_{2}]:rconst(V_{i}), \env_{2}}
\end{array}
\\ \\
\begin{array}{c}
\mylabel{T-IndexExpRead}\\
\dfrac{\begin{array}{c}
	\env_{1}, \penv \vdash e_{1}:\{\overline{F:V}\}, \env_{2}\\
	\env_{2}, \penv \vdash e_{2}:F^e, \env_{3} \;\;\;
	 F^e \lesssim F_{i}
	\end{array}}
{\env_{1}, \penv \vdash e_{1}[e_{2}]:rconst(V_{i}), \env_{3}}
\end{array}
\begin{array}{c}
\mylabel{T-CoerceClosed}\\
\dfrac{\env(id) \subtype F \;\;\; tag(F,closed)}
{\env, \penv \vdash {<}F{>} \; id:F, \env[id \mapsto reopen(F)]}
\end{array}
\begin{array}{c}
\mylabel{T-CoerceFixed}\\
\dfrac{\env(id) \subtype F \;\;\; tag(F,fixed)}
{\env, \penv \vdash {<}F{>} \; id:F, \env[id \mapsto F]}
\end{array}\\\\
\begin{array}{c}
\mylabel{T-Function}\\
\dfrac{\begin{array}{c}
	closeall(\env_{1})[\overline{id \mapsto F}], \penv[\ret \mapsto S] \vdash s, \env_{2}\\
	\env_{3} = openset(\env_{1}, frv(\mathbf{fun} \; (\overline{id{:}F}){:}S \; s)) \\
	\env_{4} = closeset(\env_{3}, fav(\mathbf{fun} \; (\overline{id{:}F}){:}S \; s))
	\end{array}}
{\env_{1}, \penv \vdash \mathbf{fun} \; (\overline{id{:}F}){:}S \; s:F_1 \times \ldots F_n \times \Void \rightarrow S, \env_{4}}
\end{array}
\begin{array}{c}
\mylabel{T-Local}\\
\dfrac{\begin{array}{c}
	\env_{1}, \penv \vdash el:P, \env_{2} \\
	P \lesssim F_1 \times \ldots \times F_n \times \Value{*} \\
	\env_{2}[\overline{id \mapsto F}], \penv \vdash s, \env_{3}
	\end{array}}
{\env_{1}, \penv \vdash \mathbf{local} \; \overline{id{:}F} = el \; \mathbf{in} \; s, (\env_{3} - \{\overline{id}\})[\overline{id \mapsto \env_2(id)}]}
\end{array}
\end{array}
\]
}
\caption{Typing Rules}
\label{fig:tablerules}
\end{figure*}

@@

Rule {\sc T-Constructor} also checks if the resulting type is {\em well-formed} with the predicate $wf$. Checking for well-formedness avoids ambiguous table types. An example of ambiguous table type is the type
$\{1:\Number, \Integer:\String, \Any:\Boolean\}$,
where indexing it with a key of type $\Integer$ could
yield either a number, a string or a boolean at runtime.

@@

\begin{figure*}[t]
{\footnotesize
\label{fig:projtypes}
\[
\begin{array}{c}
\begin{array}{c}
\mylabel{T-LocalProj}\\
\dfrac{\begin{array}{c}
	\env_{1}, \penv \vdash el:E, \env_{2}, (x,S)\\
	\env_{3} = \env_{2}[id_{1} \mapsto proj(E, 1), \ldots, id_{n} \mapsto proj(E, n)]\\
	\env_{3}, \penv[x \mapsto S] \vdash s, \env_{4} \;\;\;
	n = |\;\overline{id}\;|  
	\end{array}}
{\env_{1}, \penv \vdash \mathbf{local} \; \overline{id} = el \; \mathbf{in} \; s, (\env_{4}-\{\overline{id}\})[\overline{id \mapsto \env_2(id)}]}
\end{array}
\begin{array}{c}
\mylabel{T-ExpListProj}\\
\dfrac{\begin{array}{c}
	\env_{1}, \penv \vdash e_{i}:F_{i}, \env_{i+1} \;\;\;
	\env_{1}, \penv \vdash me:S, \env_{n+2} \;\;\; n=|\overline{e}|\\
	S = S_1 \sqcup S_2 \;\;\;m = max(|S_1|, |S_2|) \;\;\; x \notin \Pi\\
	\env_{f} = merge(\env_{1}, ..., \env_{n+2})
	\end{array}}
{\env_{1}, \penv \vdash \overline{e},me:}\\ {F_{1} \times ... \times F_{n} \times \pi_{1}^{x} \times ... \times \pi_{m}^{x}{*}, \env_{f}, (x,S)}
\end{array}\\\\
\begin{array}{c}
\mylabel{T-IfFilterProjThen}\\
\dfrac{\begin{array}{c}
	\env_{1}(id) = \pi_{i}^{x} \\ S_t = fonilpt(\Pi(x), i)\\ 
	finil(proj(\Pi(x), i)) = {\bf void} \\
	\env_{1}, \penv[x \mapsto S_t] \vdash s_{1}, \env_{2} 
	\end{array}}
{\env_{1}, \penv \vdash \mathbf{if} \; id \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}, \env_{2}}
\end{array}
\begin{array}{c}
\mylabel{T-IfFilterProjElse}\\
\dfrac{\begin{array}{c}
	\env_{1}(id) = \pi_{i}^{x} \\
	 S_e = finilpt(\Pi(x), i)\\
	fonil(proj(\Pi(x), i)) = {\bf void} \\
	\env_{1}, \penv[x \mapsto S_e] \vdash s_{2}, \env_{2} 
	\end{array}}
{\env_{1}, \penv \vdash \mathbf{if} \; id \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}, \env_{2}}
\end{array}
\begin{array}{c}
\mylabel{T-IfFilterProjBoth}\\
\dfrac{\begin{array}{c}
	\env_{1}(id) = \pi_{i}^{x} \\ S_t = fonilpt(\Pi(x), i) \; S_e = finilpt(\Pi(x), i)\\
	\env_{1}, \penv[x \mapsto S_t] \vdash s_{1}, \env_{2} \\
	\env_{1}, \penv[x \mapsto S_e] \vdash s_{2}, \env_{3} \\
	\env_{4} = join(\env_{2}, \env_{3}) 
	\end{array}}
{\env_{1}, \penv \vdash \mathbf{if} \; id \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}, \env_{4}}
\end{array}
\end{array}
\]}
\caption{Rules for Projection Types}
\end{figure*}
